---
title: "quiz 3"
author: "Andrew E. Davidson"
date: "May 24, 2015"
output: html_document
---

##Question 1
Consider the mtcars data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight as confounder. Give the adjusted estimate for the expected change in mpg comparing 8 cylinders to 4.

Using str it turns out cyl, gear, am, and carb are numeric, not factors. We'll
need to convert them
```{r}
str(mtcars)
mtcars$cyl <- factor(mtcars$cyl)
```

cyl levels are 4, 6, and 8. by default R will set the refference varriable 
(ie intercept) to the lowest level. intercept will be 4 cylinders
```{r}
summary(lm(mpg ~ cyl + wt, data=mtcars))$coefficients
c <- summary(lm(mpg ~ cyl + wt, data=mtcars))$coefficients
cyl4 <- c[1]
cyl6 <- c[2]
cyl8 <- c[3]
#cyl4 - cyl8
```
answer is cyl8 == -6.0


## question 2
Consider the mtcars data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight as a possible confounding variable. Compare the effect of 8 versus 4 cylinders on mpg for the adjusted and unadjusted by weight models. Here, adjusted means including the weight variable as a term in the regression model and unadjusted means the model without weight included. What can be said about the effect comparing 8 and 4 cylinders after looking at models with and without weight included?.

```{r}
cWithWt <- summary(lm(mpg ~ cyl + wt, data=mtcars))$coefficients
cWithWt
cWith8 <- cWithWt[3]
cWithOut <- summary(lm(mpg ~ cyl, data=mtcars))$coefficients
cWithOut
ccWithout8 <- cWithOut[3]

cWith8
ccWithout8

cWith8 - ccWithout8
```
'holding wt constant' means model with out wt


best answer
"Holding weight constant, cylinder appears to have more of an impact on mpg than if weight is disregarded."
