---
title: "quiz 4"
author: "Andrew E. Davidson"
date: "May 30, 2015"
output: html_document
---

## question 1

Consider the space shuttle data ?shuttle in the MASS library. Consider modeling the use of the autolander as the outcome (variable name use). Fit a logistic regression model with autolander (variable auto) use (labeled as "auto" 1) versus not (0) as predicted by wind sign (variable wind). Give the estimated odds ratio for autolander use comparing head winds, labeled as "head" in the variable headwind (numerator) to tail winds (denominator).

get help on libray MASS
?? MASS

If I understand the question correctly we need to change the shuttle data so that the level value auto has a numeric value of 1 and noauto has a numeric value of 0

going through earlier lecture notes. I think all we need to do is use relevel(). If my notes are correct lm() will create the dummy variables for us and set the intercept to 'no auto'

### explore shuttle data

```{r}
library(MASS)
data(shuttle)

str(shuttle$use)
levels(shuttle$use)
unique(as.numeric(shuttle$use))

tt <-shuttle$use
#yy <- relevel(tt , ref="noauto")
tt[1:5]
as.numeric(tt[1:5])
#yy[1:5]
#as.numeric(yy[1:5])

```

### solution to question 1
```{r}
# we do not need to relevel
#shuttle$use <- relevel(shuttle$use , ref="noauto")
lr <- glm(use ~ wind, family="binomial", data=shuttle)

#log odds
lr$coefficients

#odds
exp(lr$coefficients)

#Give the estimated odds ratio for autolander use comparing head winds, labeled as "head" in the variable headwind (numerator) to tail winds (denominator) ?

#went with answer for slope (ie windtail) == 0.9687
#i.e. we want odds not log odds, how ever I think we need to relevel $use
# if we relevel the answer is not in the list of solutions
exp(lr$coefficients[2])
```
