---
title: "quiz 4"
author: "Andrew E. Davidson"
date: "May 30, 2015"
output: html_document
---

## question 1

Consider the space shuttle data ?shuttle in the MASS library. Consider modeling the use of the autolander as the outcome (variable name use). Fit a logistic regression model with autolander (variable auto) use (labeled as "auto" 1) versus not (0) as predicted by wind sign (variable wind). 

Give the estimated odds ratio for autolander use comparing head winds, labeled as "head" in the variable headwind (numerator) to tail winds (denominator).?

get help on libray MASS
?? MASS

If I understand the question correctly we need to change the shuttle data so that the level value auto has a numeric value of 1 and noauto has a numeric value of 0

going through earlier lecture notes. I think all we need to do is use relevel(). If my notes are correct lm() will create the dummy variables for us and set the intercept to 'no auto'

### explore shuttle data

```{r}
library(MASS)
data(shuttle)

str(shuttle$use)
levels(shuttle$use)
unique(as.numeric(shuttle$use))
```

### solution to question 1

#### experiment 1: no relevel, give correct answer 0.969
```{r}
lr <- glm(use ~ wind, family="binomial", data=shuttle)

# log odds
round(lr$coefficients,3)

# odds
round(exp(lr$coefficients),3)

round(exp(lr$coefficients[2]),3)
```

#### experiment 2: with relevel
```{r}
shuttle$use <- relevel(shuttle$use , ref="noauto")
lr <- glm(use ~ wind, family="binomial", data=shuttle)

# log odds
round(lr$coefficients,3)

# odds
round(exp(lr$coefficients),3)

round(exp(lr$coefficients[2]),3)
```

#### experiment 3, replace $use with numeric values, 
you get the same results as if you use relevel() how ever knit has problem
```{r}

#rm(shuttle)
#data(shuttle)
#shuttle$use <- ifelse(shuttle$use == "auto",1,0)
#lr <- glm(use ~ wind, family="binomial", data=shuttle)

#log odds
#round(lr$coefficients,3)

#odds
#round(exp(lr$coefficients),3)

#Give the estimated odds ratio for autolander use comparing head winds, labeled as "head" in the variable headwind (numerator) to tail winds (denominator) ?


#went with answer for slope (ie windtail) == 0.9687
#i.e. we want odds not log odds, how ever I think we need to relevel $use
# if we relevel the answer is not in the list of solutions
#round(exp(lr$coefficients[2]),3)
```

## question 2

Consider the previous problem. Give the estimated odds ratio for autolander use comparing head winds (numerator) to tail winds (denominator) adjusting for wind strength from the variable magn.

correct answer is 0.969

```{r}
lr <- glm(use ~ wind + magn, family="binomial", data=shuttle)

#log odds
lr$coefficients

#odds
round(exp(lr$coefficients),3)

# I do not get the exact anwer> choose 0.969
```

## question 3
If you fit a logistic regression model to a binary variable, for example use of the autolander, then fit a logistic regression model for one minus the outcome (not using the autolander) what happens to the coefficients?

The outcome is binary, 0 or 1. One minus the outcome will reverse the 0's and 1: 1-0 = 1 and 1-1=0

a) The coefficients get inverted (one over their previous value).

b) The coefficients reverse their signs.

c) The coefficients change in a non-linear fashion.

d) The intercept changes sign, but the other coefficients don't.


correct answer is b. we have an "s curve", the transition from has a
positive slope and intercept is zero. When we flip, the transition has a neg slope
and intercept is 1.

```{r}
#d <- data.frame(y=c(rep(0,5),rep(1,5)), x=c(rnorm(5, mean=3, sd=0.1), rnorm(5, mean=6, sd=0.1)))
#lr <- glm(d$y ~ d$x, family="binomial")
#lr$coeffients

lr <- glm(use ~ wind, family="binomial", data=shuttle)
round(lr$coefficients,3)

shuttle$use <- relevel(shuttle$use , ref="noauto")
lr <- glm(use ~ wind, family="binomial", data=shuttle)
round(lr$coefficients,3)
```


## question 4
Consider the insect spray data InsectSprays. Fit a Poisson model using spray as a factor level. Report the estimated relative rate comapring spray A (numerator) to spray B (denominator).

```{r}
data(InsectSprays)
head(InsectSprays)
levels(InsectSprays$spray)

plr <- glm(count ~ spray, family="poisson", data=InsectSprays)
plr$coefficients
# our answer ws 0.056, how ever multple choice was -0.056???

# try using logit
# need to add one, log(count) = -inf for some values
logit <- lm(I(log(count + 1)) ~ spray, data=InsectSprays)
coef(logit)
exp(coef(logit))
```


## question 5
Consider a Poisson glm with an offset, t. So, for example, a model of the form glm(count ~ x + offset(t), family = poisson) where x is a factor variable comparing a treatment (1) to a control (0) and t is the natural log of a monitoring time. What is impact of the coefficient for x if we fit the model glm(count ~ x + offset(t2), family = poisson) where t2 <- log(10) + t? In other words, what happens to the coefficients if we change the units of the offset variable. (Note, adding log(10) on the log scale is multiplying by 10 on the original scale.)

a) The coefficient estimate is divided by 10.

b) The coefficient is subtracted by log(10).

c) The coefficient estimate is multiplied by 10.

d) The coefficient estimate is unchanged


## question 6

```{r}
x <- -5:5
y <- c(5.12, 3.93, 2.67, 1.87, 0.52, 0.08, 0.93, 2.05, 2.54, 3.87, 4.97)

# Using a knot point at 0, fit a linear model that looks like a hockey stick with two lines meeting at x=0. Include an intercept term, x and the knot point term. What is the estimated slope of the line after 0?

knots <- 0
splineTerms <- sapply(knots, function(knot) (x > knot) * (x - knot))
xMat <- cbind(1, x, splineTerms)
model <- lm(y ~ xMat - 1) # -1 means remove intercept
yhat <- predict(model)
plot(x, y, frame = FALSE, pch = 21, bg = "lightblue", cex = 2)
lines(x, yhat, col = "red", lwd = 2)

coef(model)

# answser 2.0372
```
