---
title: "README.rmd"
author: "Andrew E. Davidson"
date: "January 23, 2016"
output: 
  html_document: 
    keep_md: yes
---

# The Shinny App
This Shinny App makes it easy for users to interactively see how changes to a random variable's variance effects the shape of a Gaussian Distribution.

To run Shinny locally. Open the ui.R and server.R files using R studio. The run the following command

```{r, eval=FALSE}
library(shiny)
runApp()
```

You should see a new webspage in your browser. By default the URL will be http://127.0.0.1:3228

If you prefer a version of this shinny app has been installed on Rstudio's shiny server. You can access it at https://aedwip.shinyapps.io/developingDataProducts_CourseProject/

# The Slidy Presentation

A published version can be found at http://rpubs.com/aedwip/149315

## Changing the presetation
edit the "Developing Data Products Course Project/index.Rmd" file

Run the following to create local html and markdown files

```{r, eval=FALSE}
library(slidify)
slidify("index.Rmd")
```

## To publish changes

```{r, eval=FALSE}
publish(title = 'Developing Data Products Course Project', 'index.html', host = 'rpubs')
```
