---
title: "quiz4"
author: "Andrew E. Davidson"
date: "September 29, 2015"
output: html_document
---

## question 1

```{r load q1}
library(ElemStatLearn)
data(vowel.train)
data(vowel.test) 
```
Set the variable y to be a factor variable in both the training and test set. Then set the seed to 33833. Fit (1) a random forest predictor relating the factor variable y to the remaining variables and (2) a boosted predictor using the "gbm" method. Fit these both with the train() command in the caret package. 

What are the accuracies for the two approaches on the test data set? What is the accuracy among the test set samples where the two methods agree?

```{r q1}
vowel.train$y <- as.factor(vowel.train$y)
vowel.test$y <- as.factor(vowel.test$y)
set.seed(33833)

library(caret)
modelRF  <- train(y ~ ., data=vowel.train, method="rf")
#gbm: Generalized Boosted Regression Models
modelGBM <- train(y ~ ., data=vowel.train, method="gbm")

print(modelRF$finalModel)
print(modelGBM$finalModel)

predRF <- predict(modelRF, vowel.test)
predGBM <- predict(modelGBM, vowel.test)

cmPredRF <- confusionMatrix(predRF, vowel.test$y)
accuracyRF <- cmPredRF$overall["Accuracy"]
print(accuracyRF) # 0.6060606 

cmPredGBM <- confusionMatrix(predGBM, vowel.test$y)
accuracyGBM <- cmPredGBM$overall["Accuracy"]
print(accuracyGBM) # 0.530303

#
# What is the accuracy among the test set samples where the two methods agree?
# I think they want us to combine the two predictions? "ie ensamble"
# /courses/08_PracticalMachineLearning/025combiningPredictors/index.html
#
df <- data.frame(predRF, predGBM, y=vowel.test$y)
combModFit <- train(y ~.,method="rf",data=df)
combPred <- predict(combModFit, df)
cmCombPred <- confusionMatrix(combPred, df$y)
accuracyCmCombPred <- cmCombPred$overall["Accuracy"] # 0.6969697

```
Accuracy using RF `r  accuracyRF`
 
Accuracy using GBM `r accuracyGBM`
 
 
Accuracy combinded  `r accuracyCmCombPred`

clossed answer (our package versions do not match)
RF Accuracy = 0.6082 
GBM Accuracy = 0.5152 
Agreement Accuracy = 0.6361
