---
title: "quiz4"
author: "Andrew E. Davidson"
date: "September 29, 2015"
output: html_document
---

## question 1

```{r load q1}
library(ElemStatLearn)
data(vowel.train)
data(vowel.test) 
```
Set the variable y to be a factor variable in both the training and test set. Then set the seed to 33833. Fit (1) a random forest predictor relating the factor variable y to the remaining variables and (2) a boosted predictor using the "gbm" method. Fit these both with the train() command in the caret package. 

What are the accuracies for the two approaches on the test data set? What is the accuracy among the test set samples where the two methods agree?

```{r q1}
vowel.train$y <- as.factor(vowel.train$y)
vowel.test$y <- as.factor(vowel.test$y)
set.seed(33833)

library(caret)
modelRF  <- train(y ~ ., data=vowel.train, method="rf")
#gbm: Generalized Boosted Regression Models
modelGBM <- train(y ~ ., data=vowel.train, method="gbm")

print(modelRF$finalModel)
print(modelGBM$finalModel)

predRF <- predict(modelRF, vowel.test)
predGBM <- predict(modelGBM, vowel.test)

cmPredRF <- confusionMatrix(predRF, vowel.test$y)
accuracyRF <- cmPredRF$overall["Accuracy"]
print(accuracyRF) # 0.6060606 

cmPredGBM <- confusionMatrix(predGBM, vowel.test$y)
accuracyGBM <- cmPredGBM$overall["Accuracy"]
print(accuracyGBM) # 0.530303

#
# What is the accuracy among the test set samples where the two methods agree?
# I think they want us to combine the two predictions? "ie ensamble"
# /courses/08_PracticalMachineLearning/025combiningPredictors/index.html
#
df <- data.frame(predRF, predGBM, y=vowel.test$y)
combModFit <- train(y ~.,method="rf",data=df)
combPred <- predict(combModFit, df)
cmCombPred <- confusionMatrix(combPred, df$y)
accuracyCmCombPred <- cmCombPred$overall["Accuracy"] # 0.6969697

```
Accuracy using RF `r  accuracyRF`
 
Accuracy using GBM `r accuracyGBM`
 
 
Accuracy combinded  `r accuracyCmCombPred`

clossed answer (our package versions do not match)
RF Accuracy = 0.6082 
GBM Accuracy = 0.5152 
Agreement Accuracy = 0.6361


## Question 2
Load the Alzheimer's data using the following commands
```{r q2 load}
library(caret)
library(gbm)
set.seed(3433)
library(AppliedPredictiveModeling)
data(AlzheimerDisease)
adData = data.frame(diagnosis,predictors)
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]
training = adData[ inTrain,]
testing = adData[-inTrain,]
```
Set the seed to 62433 and predict diagnosis with all the other variables using a random forest ("rf"), boosted trees ("gbm") and linear discriminant analysis ("lda") model. Stack the predictions together using random forests ("rf"). What is the resulting accuracy on the test set? Is it better or worse than each of the individual predictions?

```{r q2 train}
set.seed(62433)
modelRF   <- train(diagnosis ~ ., data=training, method="rf")
modelGBM  <- train(diagnosis ~ ., data=training, method="gbm")
modelLDA  <- train(diagnosis ~ ., data=training, method="lda")
```

```{r q2 answer}
predRF <- predict(modelRF, testing)
predGBM <- predict(modelGBM, testing)
predLDA <- predict(modelLDA, testing)

cmPredRF <- confusionMatrix(predRF, testing$diagnosis)
accuracyRF <- cmPredRF$overall["Accuracy"] # 0.7682927 
print(accuracyRF) 

cmPredGBM <- confusionMatrix(predGBM,  testing$diagnosis)
accuracyGBM <- cmPredGBM$overall["Accuracy"] # 0.7926829 
print(accuracyGBM) 

cmPredLDA <- confusionMatrix(predLDA,  testing$diagnosis)
accuracyLDA <- cmPredLDA$overall["Accuracy"] # 0.7682927
print(accuracyLDA) 


df <- data.frame(predRF, predGBM, predLDA, diagnosis=testing$diagnosis)
combModFit <- train(diagnosis ~., method="rf", data=df)
combPred <- predict(combModFit, df)
cmCombPred <- confusionMatrix(combPred, df$diagnosis)
accuracyCmCombPred <- cmCombPred$overall["Accuracy"]  # 0.804878 
print(accuracyCmCombPred)
```

closest answer is "Stacked Accuracy: 0.88 is better than all three other methods"

## question 3

ref: 08_PracticalMachineLearning/024regularizedRegression/index.html

Load the concrete data with the commands:
```{r q3 load}
set.seed(3523)
library(AppliedPredictiveModeling)
data(concrete)
inTrain = createDataPartition(concrete$CompressiveStrength, p = 3/4)[[1]]
training = concrete[ inTrain,]
testing = concrete[-inTrain,]
```

Set the seed to 233 and fit a lasso model to predict Compressive Strength. Which variable is the last coefficient to be set to zero as the penalty increases? (Hint: it may be useful to look up ?plot.enet).


```{r q3 work}
set.seed(233)
#lasso is a regression model with regulation
modelLasso   <- train(CompressiveStrength ~ ., data=training, method="lasso")
modelLasso$finalModel
plot.enet(modelLasso$finalModel, use.color=TRUE)
```

I have no idea what the enet shows us? beta? max???. Some one on discussion
suggested the answer is cement

## question 4
ref: 08_PracticalMachineLearning/027forecasting/index.html

Load the data on the number of visitors to the instructors blog from here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/gaData.csv

```{r q4 load}
fileURL="https://d396qusza40orc.cloudfront.net/predmachlearn/gaData.csv"
dataDir="data"
dataFile <- sprintf("%s%s", dataDir, "gaData.csv")
if (!file.exists(dataDir)) { dir.create(dataDir)}
if (!file.exists(dataFile)) {
  download.file(fileURL, destfile=dataFile, method="curl")
 }

library(lubridate)  # For year() function below
dat = read.csv(dataFile)
training = dat[year(dat$date) < 2012,]
testing = dat[(year(dat$date)) > 2011,]
tstrain = ts(training$visitsTumblr)
 ```

Fit a model using the bats() function in the forecast package to the training time series. Then forecast this model for the remaining time points. For how many of the testing points is the true value within the 95% prediction interval bounds?

```{r q4 bats}
library(forecast)

modelBats <- bats(tstrain)
fcast <- forecast(modelBats, level=95)
tstest = ts(testing$visitsTumblr)
accuracy(fcast, tstest)
#fcast <- forecast(modelBats, tstest, level=95)

# understanding summary
#
# column "forcast" = fcast$mean
# the '95' is not a column it is == fcast$level
# column "hi" == fcast$upper
# column "lo" == fcast$lower
# column "point" == fcast$x
summary(fcast)
plot(fcast)
fcast
```


????


