---
title: "Practical Machine Learning Quiz 2"
author: "Andrew E. Davidson"
date: "September 22, 2015"
output: html_document
---

##Question 1
Load the Alzheimer's disease data using the commands:

```{r}
library(AppliedPredictiveModeling)
library(caret)
data(AlzheimerDisease)
```
Which of the following commands will create training and test sets with about 50% of the observations assigned to each?

```{r}
adData = data.frame(diagnosis,predictors)
trainIndex = createDataPartition(diagnosis, p = 0.50,list=FALSE)
training = adData[trainIndex,]
testing = adData[-trainIndex,]
length(diagnosis); dim(training); dim(testing)
```

## Question 2
Load the cement data using the commands:
```{r}
library(AppliedPredictiveModeling)
data(concrete)
library(caret)
set.seed(1000)
inTrain = createDataPartition(mixtures$CompressiveStrength, p = 3/4)[[1]]
training = mixtures[ inTrain,]
testing = mixtures[-inTrain,]
```

Make a histogram and confirm the SuperPlasticizer variable is skewed. Normally you might use the log transform to try to make the data more symmetric. Why would that be a poor choice for this variable?

```{r}
library(ggplot2)
qplot(concrete$Superplasticizer, geom="histogram") + ggtitle("Is SuperPlasticizer skewed?")
length(concrete$Superplasticizer);length(which(concrete$Superplasticizer == 0))
```

poor chose becuase log(0) is undefined

## Question 3
Load the Alzheimer's disease data using the commands:

```{r}
library(caret)
library(AppliedPredictiveModeling)
set.seed(3433)
data(AlzheimerDisease)
adData = data.frame(diagnosis,predictors)
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]
training = adData[ inTrain,]
testing = adData[-inTrain,]
```
Find all the predictor variables in the training set that begin with IL. Perform principal components on these variables with the preProcess() function from the caret package. Calculate the number of principal components needed to capture 80% of the variance. How many are there?

```{r}
#startsWith_IL <- names(predictors)
startsWith_IL<- predictors[,grep("^Il*", names(predictors))]

#preProc <- preProcess(log10(startsWith_IL),method="pca",pcaComp=2)
#preProc <- preProcess(log10(startsWith_IL),method="pca")
preProc <- preProcess(startsWith_IL,method="pca", thresh=0.8)
preProc
```

