---
title: "Practical Machine Learning Quiz 2"
author: "Andrew E. Davidson"
date: "September 22, 2015"
output: html_document
---

##Question 1
Load the Alzheimer's disease data using the commands:

```{r}
library(AppliedPredictiveModeling)
library(caret)
data(AlzheimerDisease)
```
Which of the following commands will create training and test sets with about 50% of the observations assigned to each?

```{r}
adData = data.frame(diagnosis,predictors)
trainIndex = createDataPartition(diagnosis, p = 0.50,list=FALSE)
training = adData[trainIndex,]
testing = adData[-trainIndex,]
length(diagnosis); dim(training); dim(testing)
```

## Question 2
Load the cement data using the commands:
```{r}
library(AppliedPredictiveModeling)
data(concrete)
library(caret)
set.seed(1000)
inTrain = createDataPartition(mixtures$CompressiveStrength, p = 3/4)[[1]]
training = mixtures[ inTrain,]
testing = mixtures[-inTrain,]
```

Make a histogram and confirm the SuperPlasticizer variable is skewed. Normally you might use the log transform to try to make the data more symmetric. Why would that be a poor choice for this variable?

```{r}
library(ggplot2)
qplot(concrete$Superplasticizer, geom="histogram") + ggtitle("Is SuperPlasticizer skewed?")
length(concrete$Superplasticizer);length(which(concrete$Superplasticizer == 0))
```

poor chose becuase log(0) is undefined
