---
title: "Practical Machine Learning Course Project"
author: "Andrew E. Davidson"
date: "September 25, 2015"
output: html_document
---

## Synopsis
aedwip

ref: http://topepo.github.io/caret/preprocess.html

## Data processing

Down load the data and documentation
```{r load training and testing data }
fileURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
fileName <- "pml-training.csv"
dataDir <- "data/"
trainingDataFile <- sprintf("%s%s", dataDir, fileName)
if (!file.exists(dataDir)) { dir.create(dataDir)}
if (!file.exists(trainingDataFile)) {
  download.file(fileURL, destfile=trainingDataFile, method="curl")
  dateDownLoaded <- date()
  write(dateDownLoaded, file=sprintf("%s%s", dataDir, "dateDownLoaded.txt"))
}

fileURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
fileName <- "pml-testing.csv"
testingDataFile <- sprintf("%s%s", dataDir, fileName)
if (!file.exists(testingDataFile)) {
  download.file(fileURL, destfile=testingDataFile, method="curl")
}
```

Down load the data releated documentation if needed
```{r down load documentation}
fileURL <- "http://groupware.les.inf.puc-rio.br/har"
fileName <- "har.html"
docFile <- sprintf("%s%s", dataDir, fileName)
if (!file.exists(docFile)) {
  download.file(fileURL, destfile=docFile, method="curl")
}
```

Load data into memory and remove tempory variables. We'll temporarly turn off warnings about missing objects not found

```{r load data into memory}
library(data.table)
if (!exists("rawTrainTable")){
    csv <- read.csv(trainingDataFile, header=TRUE, fill=TRUE)
    rawTrainTable <- data.table(csv)
}

if (!exists("rawTestTable")){
    csv <- read.csv(testingDataFile, header=TRUE, fill=TRUE)
    rawTestTable <- data.table(csv)
}
options(warn=-1) 
rm(csv, dataDir, dataFileZip, dateDownLoaded, testingDataFile,trainingDataFile, fileURL, docFile, fileName)
options(warn=0) 
```


## Exploring the data
Documentation states accelorometers are placed on the subjects forearm, arm,
belt, and on the dumbbell. It appears there is a lot of summary releated data
for each sample. We should probably just work the observed data

```{r explore data lots of junk}
grep("roll_forearm", names(rawTrainTable), value=TRUE)
```

How does time data work? 
```{r exploring time}
grep("time", names(rawTrainTable), value=TRUE)

agg <- rawTrainTable[user_name=="carlitos", sum(raw_timestamp_part_2), by=raw_timestamp_part_1]
dim(agg)
agg[1:10,]

```

## Pre processing and selecting data
We have a lot of data. There is a good chance a lot of it is not useful. We'll 
want to reduce the number of featues we use if possible. Our model will train
faster and be easier to understand.In general we do not want to edit the raw data we down loaded. Instead we create
a new clean data set. 

The first thing we want to get rid of the first 7 columns.
"X" appears to identify each observation. If we need to know who the subject
our model will not generalize well. The other variable have to do with time stamp
and 'binning' samples based on time. If our model does not work well we can
consider using them in a future analysis.

```{r pre processing and selcting data remove first 7 columns}
rmCols <- names(rawTrainTable)[1:7]
rmCols
t1 <- rawTrainTable[,!rmCols, with=FALSE]
rm(rmCols)
```

Next we want to find and remove any features that are missing most of their data
```{r remove features with lots of missing data}
n <- nrow(t1) 
lotsOfNA <- t1[, lapply(.SD, function(x) (sum(is.na(x), na.rm=TRUE) / n)) > 0.5]
rmColNames <- names(t1)[lotsOfNA]
t2 <- t1[,!rmColNames, with=FALSE]
rm(t1, lotsOfNA, rmColNames)
```

Check to see if any of the factors haveZero- and Near Zero-Variance. If remove
them, they will not be good predictors

```{r look for near zero variance factors}
nzv <- nearZeroVar(t2, saveMetrics= TRUE)
nearZeroCols <- nzv[nzv$nzv,4]
numberOfNearZeroFeatures <- sum(nearZeroCols)
rmColNames <- names(t2)[nzv[,4]]
t3 <- t2[,!rmColNames, with=FALSE]
rm(t2, nearZeroCols, rmColNames)
```

##Random forest
our label is 'classe'


```{r train model}
library(caret)
#rpart is regression classifier
#modFit <- train(classe ~ .,method="rpart",data=trainTable)
#print(modFit$finalModel)

# rf is random forest
#system.time(modFit <- train(classe ~ .,data=t2, method="rf", prox=TRUE))
#modFit

```

