---
title: "Statistical Inference Course Project"
author: "Andrew E. Davidson"
date: "February 22, 2015"
output: pdf_document
---

My R Markdown file borrows heavily from 
https://github.com/bcaffo/courses/blob/master/06_StatisticalInference/03_01_TwoGroupIntervals

---

### 1. Load the ToothGrowth data and perform some basic exploratory data analyses 

In RStudio run help(ToothGrowth) to see a description of the data set.

Description

The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

```{r}
str(ToothGrowth)

head(ToothGrowth)

nrow(ToothGrowth)
nrow(na.omit(ToothGrowth) )


```

The following graph suggest their is not much difference in tooth growth based on treatment

```{r}
library(ggplot2)
vc <- ToothGrowth$len[ToothGrowth$supp == "VC"]
oj <- ToothGrowth$len[ToothGrowth$supp == "OJ"]
x <- 1:length(vc)
plotData <- data.frame(oj, vc, x)
pp <- ggplot(plotData, aes(x, title="len by treatment")) 
pp + geom_line(aes(y=oj, colour="oj")) + geom_line(aes(y=vc, colour="vc"))
```
---

### 2. Provide a basic summary of the data.

```{r}
summary(ToothGrowth)

```
---

### 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. 

We'll use "Two Group intervals Statistical Inference"

First create 6 list, one for each combination of treament and dose
```{r}
vc05 <- ToothGrowth$len[ToothGrowth$supp == "VC" & ToothGrowth$dose == 0.5]
vc1 <- ToothGrowth$len[ToothGrowth$supp == "VC" & ToothGrowth$dose == 1.0]
vc2 <- ToothGrowth$len[ToothGrowth$supp == "VC" & ToothGrowth$dose == 2.0]

oj05 <- ToothGrowth$len[ToothGrowth$supp == "OJ" & ToothGrowth$dose == 0.5]
oj1 <- ToothGrowth$len[ToothGrowth$supp == "OJ" & ToothGrowth$dose == 1.0]
oj2 <- ToothGrowth$len[ToothGrowth$supp == "OJ" & ToothGrowth$dose == 2.0]

```

Now for each does we'll compare the VC vs OJ

```{r}
mean(vc05) - mean(oj05)
t.test(vc05, oj05, paired = FALSE, var.equal = TRUE)$conf
```

```{r}
mean(vc1) - mean(oj1)
t.test(vc1, oj1, paired = FALSE, var.equal = TRUE)$conf
```

```{r}
mean(vc2) - mean(oj2)
t.test(vc2, oj2, paired = FALSE, var.equal = TRUE)$conf
```


Lets see if we find a differece between treaments when we ignore the dose

```{r}
vc <- ToothGrowth$len[ToothGrowth$supp == "VC"]
oj <- ToothGrowth$len[ToothGrowth$supp == "OJ"]

mean(vc) - mean(oj)
t.test(vc, oj, paired = FALSE, var.equal = TRUE)$conf
```

---

### 4. State your conclusions and the assumptions needed for your conclusions. 

regardless of the size of the dose guinea pigs recevied, there does not appear to be a statistically signifigant difference in the amount of tooth growth.

The only assumption I made was groups where independent so that we could not use a paired t test

